---
title: "Designing glycoform panels for biosynthetic-pathway inference"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Designing glycoform panels for biosynthetic-pathway inference}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

## The design question

A measured composition of $m$ glycoform classes has at most $m-1$ independent
directions because the proportions sum to one. That dimension bound is
necessary but not sufficient to recover $k$ specified biosynthetic rate
directions. Their local response patterns must also be linearly independent,
well conditioned, and large enough relative to assay noise. Finally, the
forward model must adequately represent the observed composition.

`glycoPathDesign` keeps these requirements separate.

## Define the biochemical claim

```{r}
library(glycoPathDesign)
pathway <- canonical_pathway()
pathway
```

Each edge belongs to an effective `rate_class`. Edges with the same class share
one parameter. This grouping defines the biological distinctions being tested;
changing it changes the inverse problem.

## Audit a proposed measurement

```{r}
panel <- c("M9", "M8", "M5", "M5Gn", "M3Gn",
           "M3Gn2", "M3Gn3", "G1", "G2S1")
audit <- check_panel(pathway, panel, target_sd = 0.25, assay_cv = 0.01)
audit
```

The log-rate Jacobian $J$ contains the local change in every measured
composition component caused by each log effective rate. Full column rank means
that those rate directions have distinct local signatures. The condition number
is the ratio between the strongest and weakest singular directions. The
amplification factor $g$ incorporates the stated compositional error model and
gives the worst-direction relation

$$
\mathrm{SD}(\log r) \approx g\,\mathrm{CV}(y).
$$

Thus the maximum admissible assay CV for a target log-rate SD is
`target_sd / g`.

## Check more than one operating point

Local rank and conditioning can change with the effective rates. A design
should therefore be tested over a scientifically plausible set of operating
points.

```{r}
robust <- check_operating_points(pathway, panel, n = 25, log_rate_sd = 0.5, seed = 1)
robust
```

The sampling distribution is part of the scientific assumption and should be
reported. A random robustness sample is not an exhaustive global proof.

## Search candidate panels

The following deliberately restricts the candidate set so that the vignette is
quick to build.

```{r}
candidate_classes <- c(panel, "G2", "G1F")
sets <- enumerate_panels(pathway, 9, candidates = candidate_classes,
                         max_panels = 200, seed = 1)
sets
optimise_panels(sets, n = 5)
```

When all combinations are evaluated, the result can establish panel minimality
for that candidate set, graph, operating point, and numerical rank tolerance.
When combinations are sampled, the object explicitly records `sampled = TRUE`.

## Fit is a separate requirement

```{r}
observed <- read.csv(system.file("extdata", "example_observed.csv",
                                 package = "glycoPathDesign"))
fit <- fit_pathway(pathway, observed, n_starts = 2)
fit
```

A good fit shows that the pathway has capacity to represent the composition. It
does not by itself show that individual effective rate classes are recoverable;
the audit at the fitted operating point provides that second assessment.
